\bf Forecast Reconciliation: A geometric view with new insights on bias correction.

### 2019-09-04 Hierarchical-forecast-reconciliation/HierarchicalForecastReconciliation.tex
Final Revision

	\begin{abstract} 
		A geometric interpretation is developed for so-called {\em reconciliation} methodologies which are used to forecast time series that adhere to known linear constraints.  In particular, a general framework is established nesting many existing popular reconciliation methods within the class of {\em projections}.  This interpretation facilitates the derivation of novel results that explain why and how reconciliation via projection is guaranteed to improve forecasts with respect to a specific class of loss functions.  This result is later demonstrated empirically.  The geometric interpretation is also used to provide a new proof that forecast reconciliation results in unbiased forecasts as long as the initial so-called {\em base} forecasts are also unbiased.  Approaches for dealing with biased base forecasts are proposed and explored in an extensive empirical study on Australian tourism demand.  Overall, the method of bias-correcting before carrying out reconciliation is shown to outperform alternatives that only bias-correct or only reconcile forecasts.

@@

    \begin{equation}
    \bm{y}_t = \bm{Sb}_t,
    \end{equation}
    where $\bm{S}$ is an $n \times m$ constant matrix that encodes the aggregation constraints.

@@

	\begin{theo}[Distance reducing property]\label{th:distred}
		If $\tilde{\bm{y}}_{t+h|t}=\bm{S}\bm{G}\hat{\bm{y}}_{t+h|t}$, where $\bm{G}$ is such that $\bm{S}\bm{G}$ is an orthogonal (in the Euclidean sense) projection onto $\mathfrak{s}$ and let $\|\bm{v}\|$ be the $L_2$ norm (in the Euclidean sense) of vector $\bm{v}$ then:
		\begin{equation}
		\|(\bm{y}_{t+h}-\tilde{\bm{y}}_{t+h|t})\|\le\|(\bm{y}_{t+h}-\hat{\bm{y}}_{t+h|t})\|.
		\end{equation}
	\end{theo}
	\begin{proof}
		Since, $\bm{y}_{t+h|t},\tilde{\bm{y}}_{t+h|t}\in\mathfrak{s}$ and since the projection is orthogonal, by Pythagoras' theorem
		\begin{equation} \|(\bm{y}_{t+h}-\hat{\bm{y}}_{t+h|t})\|^2=\|(\bm{y}_{t+h}-\tilde{\bm{y}}_{t+h|t})\|^2+\|(\tilde{\bm{y}}_{t+h|t}-\hat{\bm{y}}_{t+h|t})\|^2.
		\end{equation}

		Since $\|(\tilde{\bm{y}}_{t+h|t}-\hat{\bm{y}}_{t+h|t})\|^2\ge 0$ this implies,
		\begin{equation} \|(\bm{y}_{t+h}-\hat{\bm{y}}_{t+h|t})\|^2\ge\|(\bm{y}_{t+h}-\tilde{\bm{y}}_{t+h|t})\|^2.
		\end{equation}
		with equality only holding when $\tilde{\bm{y}}_{t+h|t}=\hat{\bm{y}}_{t+h|t}$.  Taking the square root of both sides proves the desired result.
	\end{proof}

@@

	\begin{align}
	\tilde{\bm{y}}^*&=\bm{W}^{-1/2}\tilde{\bm{y}}\\&=\bm{W}^{-1/2}\bm{S}\left(\bm{S}'\bm{W}^{-1}\bm{S}\right)^{-1}\bm{S}'\bm{W}^{-1}\hat{\bm{y}}
	\\&=\bm{S}^*\left(\bm{S}^{*'}\bm{S}^*\right)^{-1}\bm{S}^{*'}\hat{\bm{y}^*}.
	\end{align}

@@

	\begin{theo}[General distance reducing property]\label{th:gdistred}
        If $\tilde{\bm{y}}_{t+h|t}=\bm{S}\bm{G}\hat{\bm{y}}_{t+h|t}$, where $\bm{G}$ is such that $\bm{S}\bm{G}$ is an orthogonal (in the generalised Euclidean sense) projection onto $\mathfrak{s}$ then:
		\begin{equation}
		\|(\bm{y}_{t+h}-\tilde{\bm{y}}_{t+h|t})\|_{{\bm W}^{-1}}\le\|(\bm{y}_{t+h}-\hat{\bm{y}}_{t+h|t})\|_{{\bm W}^{-1}}.
		\end{equation}
	\end{theo}
	\begin{proof}
    The proof is identical to the proof for Theorem~\ref{th:distred} but relies on the generalised Pythagorean Theorem (applicable to Generalised Euclidean space) rather than the Pythagorean Theorem.
    \end{proof}

@@

    This result has implications in practice. The top-down method \citep{Gross1990} has
	\begin{equation}\label{eq:top-downG}
	\bm{G}=\begin{pmatrix}
	\bm{p} & \bm{0}_{(m \times n-1)}
	\end{pmatrix}
	\end{equation}
    where $\bm{p} = (p_1,\dots,p_m)'$ is an $m$-dimensional vector consisting a set of proportions used to disaggregate the top-level forecast.  In this case it can be verified that $\bm{SG}$ is idempotent, i,e. $\bm{SGSG}=\bm{SG}$ and therefore $\bm{SG}$ is a projection matrix.  However the image of this projection is not an $m$-dimensional subspace but a $1$-dimensional subspace.  As such, top-down reconciliation produces biased forecasts even when the base forecasts are unbiased.

@@

	Finally, it is often stated that an assumption required to prove the unbiasedness preserving property is that $\bm{SGS}=\bm{S}$ or alternatively that $\bm{GS}=\bm{I}$.  Both of these conditions are equivalent to assuming that $\bm{SG}$ is a projection matrix. However, problems arise when viewing the preservation of unbiasedness through the prism of imposing a constraint $\bm{GS}=\bm{I}$. This thinking suggests that a way to deal with biased forecasts is to select $\bm{G}$ in an unconstrained manner.  However, equipped with a geometric understanding of the problem, we would advise against this approach.  The constraint $\bm{GS}=\bm{I}$  is not just about bias. Dropping the constraint compromises all of the attractive properties of projections.  Opening the door to reconciliation methods that change already coherent base forecasts would seem to suggest an increase in the variability of the forecast.  This seems particularly perverse when the motivation for using a biased method in the first place is to reduce variance.

@@

	Our own solution to dealing with biased forecasts is to bias correct {\em before} reconciliation.  In many cases the method for bias correction will be context specific.  For instance, in  our empirical study in Section~\ref{sec:EmpStudy} we consider a scenario where bias is induced via taking either a log transformation or a Box-Cox transformation before modelling.  In this well-known case a number of bias correction methods exist based on Taylor expansions. %\textcolor{red}{TODO: Read this again after the empirical application has been completed. I am saying this because my understanding is that we use the Guererro method to choose the lambda for the BoxCox transformation. Then we back transform the forecasts using the same lambda once with bias correcting using the factor that comes out of a Taylor series expansion around $\mu$ (programmed in fable/forecast package) and once without it hence looking at medians and not means - hence biased base forecasts. If my understanding here is correct I am not sure we can say we are using the Guererro method to bias correct.}

@@

	%\textcolor{red}{GA: I will rewrite/review this when we decide where we put all our results - I am referring to the projection results - currently I think it is best that everything goes here and we make a comment at the end of Section 3.3 guiding the reader to the results here. This will make the empirical application a bit broader not only about bias correction, i.e., both about showing how projections work and how bias correction works.}
	Using an empirical application to forecast Australian domestic tourism flows, we illustrate the usefulness of projection based reconciliation in practice. Previous studies have found that reconciliation improves point forecast accuracy in domestic tourism flows for Australia \citep[see for example][]{Athanasopoulos2009, Hyndman2011,WicEtAl2019}. However, our motivation in this study is twofold.  First, we demonstrate the implications of Theorem~\ref{th:distred} by comparing reconciled and base forecasts. In contrast to previous studies we consider individual periods rather than computing averages over a rolling window.  Second, we demonstrate how the bias correction methods discussed in the previous section along with the projection-based reconciliation help to improve forecast accuracy.

@@

\begin{equation} \label{eq:BoxCox_BT_biasadj}
	 \hat{y}_{t+h|t} =
	\begin{cases}
	\exp({\hat{w}_{t+h|t}})[1+\frac{\sigma_h^2}{2}] & \text{if}~\lambda = 0, \\
	(\lambda \hat{w}_{t+h|t} + 1)^{1/\lambda}[1 + \frac{\sigma_h^2(1-\lambda)}{2(\lambda \hat{w}_{t+h|t} + 1)^2}]       & \text{if}~\lambda \ne 0,
	\end{cases}
	\end{equation}
where, $\hat{w}_{t+h|t}$ is the $h$-step-ahead forecast from the Box-Cox transformed series and $\sigma_h^2$ is the variance of $\hat{w}_{t+h|t}$. Using the mean of the forecast distribution returns bias adjusted base forecasts compared to the simple back-transformation of \eqref{eq:BoxCox_back-transformation}. We refer to this as Method-1 in the results that follow. The second scenario of bias adjustment we explore is using the in-sample forecast error mean of the biased forecasts to adjust the out-of-sample forecasts. We refer to this as Method-2 in the results that follow.

@@

\begin{table}
	\caption {Average {MSE($\times 10^3$)} of $1$-step-ahead point forecasts from log and Box-Cox transformed series. Biased denotes forecasts from simply reversing the transformation via \eqref{eq:BoxCox_back-transformation}. Unbiased(Method-1) performs bias adjustment via a Taylor series expansion as shown in \eqref{eq:BoxCox_BT_biasadj} whereas Unbiased(Method-2) bias adjusts by subtracting the in-sample forecast error mean.}
	\label{tab:Results_MSE}
	\centering
		\begin{tabular}{lcccccc}
			\toprule
			\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Log Transformation} &
			\multicolumn{3}{c}{Box-Cox Transformation}\\
			\cmidrule(lr){2-4} \cmidrule(lr){5-7}
			Method & Biased & Unbiased & Unbiased & Biased & Unbiased & Unbiased \\
			& & (Method-1) & (Method-2) && (Method-1) &(Method-2)\\ \midrule
			Base      & 4.47 & 4.43 & 4.51 & 4.59 & 4.52 & 4.78\\
			OLS       & 4.34 & 4.31 & 4.36 & 4.46 & 4.40 & 4.63\\
			MinT      & 4.32 & \textbf{4.16} & 4.61 & 4.29 & \textbf{4.17} & 4.54\\
			WLS       & 4.82 & 4.38 & 5.43 & 4.77 & 4.42 & 5.23\\
			Bottom-up & 6.36 & 5.26 & 8.06 & 6.21 & 5.31 & 7.46\\
			\bottomrule
		\end{tabular}	
\end{table}

@@

	\section{Appendix}
	%\thisfloatpagestyle{empty}
	\subsection{Proof $\bm{S}\bm{G}\bm{S}=\bm{S}$ implies $\bm{S}\bm{G}$ is a projection}
	Here we establish that if $\bm{S}\bm{G}$ is a projection onto the linear subspace spanned by $\bm{S}$ then $\bm{S}\bm{G}\bm{S}=\bm{S}$.  We also prove that the converse holds, namely that if the condition $\bm{S}\bm{G}\bm{S}=\bm{S}$ holds then $\bm{S}\bm{G}$ must be a projection onto the linear subspace spanned by $\bm{S}$.

	To establish the first statement let $\bm{s}_j$ be the $j^{th}$ column of $\bm{S}$.  Since by definition, $\bm{s}_j$ lies in $\mathfrak{s}$, it must hold that $\bm{S}\bm{G}\bm{s}_j=\bm{s}_j$.  Stacking these vectors horizontally
	\begin{align}
	\bm{S}\bm{G}\bm{S}&=\begin{pmatrix}
	\bm{S}\bm{G}\bm{s}_1, & \bm{S}\bm{G}\bm{s}_2, & \cdots & \bm{S}\bm{G}\bm{s}_m
	\end{pmatrix}\\
	&=\begin{pmatrix}
	\bm{s}_1, & \bm{s}_2, & \cdots & \bm{s}_m
	\end{pmatrix}\\
	&=\bm{S}
	\end{align}

>>>>

	\begin{abstract}
\spacingset{1.45}
        A geometric interpretation is developed for so-called {\em reconciliation} methodologies used to forecast time series that adhere to known linear constraints.  In particular, a general framework is established nesting many existing popular reconciliation methods within the class of {\em projections}.  This interpretation facilitates the derivation of novel results that explain why and how reconciliation via projection is guaranteed to improve forecast accuracy with respect to a specific class of loss functions.  The result is also demonstrated empirically.  The geometric interpretation is further used to provide a new proof that forecast reconciliation results in unbiased forecasts as long as the initial so-called {\em base} forecasts are also unbiased.  Approaches for dealing with biased base forecasts are proposed and explored in an extensive empirical study on Australian tourism flows.  Overall, the method of bias-correcting before carrying out reconciliation is shown to outperform alternatives that only bias-correct or only reconcile forecasts.

@@

    \begin{equation*}
    \bm{y}_t = \bm{Sb}_t,
    \end{equation*}
    where $\bm{S}$ is an $n \times m$ constant matrix that encodes the aggregation constraints.

@@

	\begin{theo}[Distance reducing property]\label{th:distred}
		If $\tilde{\bm{y}}_{t+h|t}=\bm{S}\bm{G}\hat{\bm{y}}_{t+h|t}$, where $\bm{G}$ is such that $\bm{S}\bm{G}$ is an orthogonal (in the Euclidean sense) projection onto $\mathfrak{s}$ and let $\|\bm{v}\|$ be the $L_2$ norm (in the Euclidean sense) of vector $\bm{v}$ then:
		\begin{equation*}
		\|(\bm{y}_{t+h}-\tilde{\bm{y}}_{t+h|t})\|\le\|(\bm{y}_{t+h}-\hat{\bm{y}}_{t+h|t})\|.
		\end{equation*}
	\end{theo}
	\begin{proof}
		Since, $\bm{y}_{t+h|t},\tilde{\bm{y}}_{t+h|t}\in\mathfrak{s}$ and since the projection is orthogonal, by Pythagoras' theorem
		\begin{equation*} \|(\bm{y}_{t+h}-\hat{\bm{y}}_{t+h|t})\|^2=\|(\bm{y}_{t+h}-\tilde{\bm{y}}_{t+h|t})\|^2+\|(\tilde{\bm{y}}_{t+h|t}-\hat{\bm{y}}_{t+h|t})\|^2.
		\end{equation*}

		Since $\|(\tilde{\bm{y}}_{t+h|t}-\hat{\bm{y}}_{t+h|t})\|^2\ge 0$ this implies,
		\begin{equation*} \|(\bm{y}_{t+h}-\hat{\bm{y}}_{t+h|t})\|^2\ge\|(\bm{y}_{t+h}-\tilde{\bm{y}}_{t+h|t})\|^2.
		\end{equation*}
		with equality only holding when $\tilde{\bm{y}}_{t+h|t}=\hat{\bm{y}}_{t+h|t}$.  Taking the square root of both sides proves the desired result.
	\end{proof}

@@

	\begin{align*}
	\tilde{\bm{y}}^*&=\bm{W}^{-1/2}\tilde{\bm{y}}\\&=\bm{W}^{-1/2}\bm{S}\left(\bm{S}'\bm{W}^{-1}\bm{S}\right)^{-1}\bm{S}'\bm{W}^{-1}\hat{\bm{y}}
	\\&=\bm{S}^*\left(\bm{S}^{*'}\bm{S}^*\right)^{-1}\bm{S}^{*'}\hat{\bm{y}^*}.
	\end{align*}

@@

	\begin{theo}[General distance reducing property]\label{th:gdistred}
        If $\tilde{\bm{y}}_{t+h|t}=\bm{S}\bm{G}\hat{\bm{y}}_{t+h|t}$, where $\bm{G}$ is such that $\bm{S}\bm{G}$ is an orthogonal (in the generalised Euclidean sense) projection onto $\mathfrak{s}$ then:
		\begin{equation*}
		\|(\bm{y}_{t+h}-\tilde{\bm{y}}_{t+h|t})\|_{{\bm W}^{-1}}\le\|(\bm{y}_{t+h}-\hat{\bm{y}}_{t+h|t})\|_{{\bm W}^{-1}}.
		\end{equation*}
	\end{theo}
	\begin{proof}
    The proof is identical to the proof for Theorem~\ref{th:distred} but relies on the generalised Pythagorean Theorem (applicable to Generalised Euclidean space) rather than the Pythagorean Theorem.
    \end{proof}

@@

    This result has implications in practice. The top-down method \citep{Gross1990} has
	\begin{equation*}
	\bm{G}=\begin{pmatrix}
	\bm{p} & \bm{0}_{(m \times n-1)}
	\end{pmatrix}
	\end{equation*}
    where $\bm{p} = (p_1,\dots,p_m)'$ is an $m$-dimensional vector consisting a set of proportions used to disaggregate the top-level forecast.  In this case it can be verified that $\bm{SG}$ is idempotent, i,e. $\bm{SGSG}=\bm{SG}$ and therefore $\bm{SG}$ is a projection matrix.  However the image of this projection is not an $m$-dimensional subspace but a $1$-dimensional subspace.  As such, top-down reconciliation produces biased forecasts even when the base forecasts are unbiased.

@@

	Finally, it is often stated that an assumption required to prove the unbiasedness preserving property is that $\bm{SGS}=\bm{S}$ or alternatively that $\bm{GS}=\bm{I}$.  Both of these conditions are equivalent to assuming that $\bm{SG}$ is a projection matrix (see Section \ref{app:projection-proof} in the Appendix for a proof).\todo{Rob can you please read this carefully and see if you are happy with the following statements about bias correction.} However, problems arise when viewing the preservation of unbiasedness through the prism of imposing a constraint $\bm{GS}=\bm{I}$. This thinking suggests that a way to deal with biased forecasts is to select $\bm{G}$ in an unconstrained manner.  However, equipped with a geometric understanding of the problem, we would advise against this approach.  The constraint $\bm{GS}=\bm{I}$  is not just about bias. Dropping the constraint compromises all of the attractive properties of projections.  Opening the door to reconciliation methods that change already coherent base forecasts would seem to suggest an increase in the variability of the forecast.  This seems particularly perverse when the motivation for using a biased method in the first place is to reduce variance.

@@

	Our own solution to dealing with biased forecasts is to bias correct {\em before} reconciliation.  In many cases the method for bias correction will be context specific.  For instance, in  our empirical study in Section~\ref{sec:EmpStudy} we consider a scenario where bias is induced via taking either a log transformation or a Box-Cox transformation before modelling.  In this well-known case a number of bias correction methods exist based on Taylor expansions.

@@

	Using an empirical application to forecast Australian domestic tourism flows, we illustrate the usefulness of projection based reconciliation in practice. Previous studies have found that reconciliation improves point forecast accuracy in domestic tourism flows for Australia \citep[see for example][]{Athanasopoulos2009, Hyndman2011,WicEtAl2019}. However, our motivation in this study is twofold.  First, we demonstrate the implications of Theorem~\ref{th:distred} by comparing reconciled and base forecasts. In contrast to previous studies we consider individual periods rather than computing averages over a rolling window.  Second, we demonstrate how the bias correction methods discussed in the previous section along with the projection-based reconciliation help to improve forecast accuracy.

@@

\begin{equation} \label{eq:BoxCox_BT_biasadj}
	 \hat{y}_{t+h|t} =
	\begin{cases}
	\exp({\hat{w}_{t+h|t}})[1+\frac{\sigma_h^2}{2}] & \text{if}~\lambda = 0, \\
	(\lambda \hat{w}_{t+h|t} + 1)^{1/\lambda}[1 + \frac{\sigma_h^2(1-\lambda)}{2(\lambda \hat{w}_{t+h|t} + 1)^2}]       & \text{if}~\lambda \ne 0,
	\end{cases}
	\end{equation}
where, $\hat{w}_{t+h|t}$ is the $h$-step-ahead forecast from the Box-Cox transformed series and $\sigma_h^2$ is the variance of $\hat{w}_{t+h|t}$. Using the mean of the forecast distribution returns bias adjusted base forecasts compared to the simple back-transformation of Eq.~\eqref{eq:BoxCox_back-transformation}. We refer to this as ``Method-1'' in the results that follow. The second scenario of bias adjustment we explore is using the in-sample forecast error mean of the biased forecasts to adjust the out-of-sample forecasts. We refer to this as ``Method-2'' in the results that follow.

@@

\begin{table}
	\caption {Average {MSE($\times 10^3$)} of $1$-step-ahead point forecasts from log and Box-Cox transformed series. Biased denotes forecasts from simply reversing the transformation via Eq.~\eqref{eq:BoxCox_back-transformation}. Unbiased(Method-1) performs bias adjustment via a Taylor series expansion as shown in Eq.~\eqref{eq:BoxCox_BT_biasadj} whereas Unbiased(Method-2) bias adjusts by subtracting the in-sample forecast error mean.}
	\label{tab:Results_MSE}
	\centering
		\begin{tabular}{lcccccc}
			\toprule
			\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Log Transformation} &
			\multicolumn{3}{c}{Box-Cox Transformation}\\
			\cmidrule(lr){2-4} \cmidrule(lr){5-7}
			Method & Biased & Unbiased & Unbiased & Biased & Unbiased & Unbiased \\
			& & (Method-1) & (Method-2) && (Method-1) &(Method-2)\\ \midrule
			Base      & 4.47 & 4.43 & 4.51 & 4.59 & 4.52 & 4.78\\
			OLS       & 4.34 & 4.31 & 4.36 & 4.46 & 4.40 & 4.63\\
			MinT      & 4.32 & \textbf{4.16} & 4.61 & 4.29 & \textbf{4.17} & 4.54\\
			WLS       & 4.82 & 4.38 & 5.43 & 4.77 & 4.42 & 5.23\\
			Bottom-up & 6.36 & 5.26 & 8.06 & 6.21 & 5.31 & 7.46\\
			\bottomrule
		\end{tabular}	
\end{table}

@@

\clearpage
	\section{Appendix} 
	%\thisfloatpagestyle{empty}
	\subsection{Proof $\bm{S}\bm{G}\bm{S}=\bm{S}$ implies $\bm{S}\bm{G}$ is a projection} \label{app:projection-proof}
	Here we establish that if $\bm{S}\bm{G}$ is a projection onto the linear subspace spanned by $\bm{S}$ then $\bm{S}\bm{G}\bm{S}=\bm{S}$.  We also prove that the converse holds, namely that if the condition $\bm{S}\bm{G}\bm{S}=\bm{S}$ holds then $\bm{S}\bm{G}$ must be a projection onto the linear subspace spanned by $\bm{S}$.

	To establish the first statement let $\bm{s}_j$ be the $j^{th}$ column of $\bm{S}$.  Since by definition, $\bm{s}_j$ lies in $\mathfrak{s}$, it must hold that $\bm{S}\bm{G}\bm{s}_j=\bm{s}_j$.  Stacking these vectors horizontally
	\begin{align*}
	\bm{S}\bm{G}\bm{S}&=\begin{pmatrix}
	\bm{S}\bm{G}\bm{s}_1, & \bm{S}\bm{G}\bm{s}_2, & \cdots & \bm{S}\bm{G}\bm{s}_m
	\end{pmatrix}\\
	&=\begin{pmatrix}
	\bm{s}_1, & \bm{s}_2, & \cdots & \bm{s}_m
	\end{pmatrix}\\
	&=\bm{S}
	\end{align*}
